The LBRY Trending Algorithm

### 2021-12-15 tex/ms.tex
Figure stuff
More writing

\title{The LBRY Trending Algorithm}
\author{Brendon J. Brewer}
\date{}

@@

\section{Introduction}
LBRY \citep{li2020lbry} needs a trending algorithm to rank claims by `recent popularity',
preferably using on-chain information such as changes in support. An
intuitive idea is to have increases and decreases in overall support cause
a `spike' in the trending score, which then fades away over time
(assuming no further changes occur). This idea has formed the basis of all the
implementations that I have written.

@@

In 2021, Jack Robison rewrote a lot of the hub code to improve performance
and to ensure hub data integrity when a chain reorganisation occurs. The
backend now uses leveldb to store transaction and claim information, and
Elasticsearch to store claim information in a suitable form for clients to
query. Leveldb will be replaced with rocksdb soon.
As part of these changes, trending needed to be made more efficient. This
meant we made some modifications to it, some of which are pretty neat
in my opinion.

@@

% TODO: Consider idea of symmetrising by using both old and new LBC values,
% as this may make reorg safety occur naturally without any extra effort

@@

\section{The `Squashed' Grade}
Suppose $x$ is a huge number, where double precision overflow might occur if
we were to use it in a calculation (this happens at around $10^{308}$).
If $x$ is positive, we could instead deal with logarithms. Then, to add
two huge numbers without overflow, we could use
\begin{align}
\log(x + y)  &= \log(\exp(\log x) + \exp(\log y)) \\
             &= \logsumexp(\log x, \log y),
\end{align}
where $\logsumexp()$ is given by
\begin{align}
\logsumexp(a, b) &= \max(a, b)
        + \log\left[\exp(a - \max(a, b)) + \exp(b - \max(a, b))\right].
\end{align}
Subtracting the maximum before doing the exponentiation prevents overflow.
Similarly, differences could be computed on the log scale. Assuming $x>y$,
we can use the following method:
\begin{align}
\log(x - y)  &= \log(\exp(\log x) - \exp(\log y)) \\
             &= \logdiffexp(\log x, \log y),
\end{align}
where $\logdiffexp$ is given by
\begin{align}
\logdiffexp(a, b) &= a
        + \log\left[1 - \exp(b-a)\right]
\end{align}
for $a > b$.

@@

@@

\section{Acknowledgements}
BJB would like to thank Kevin Knuth (SUNY Albany) for inspiration.

>>>>

\title{The LBRY Trending Algorithm}
\author{Brendon J. Brewer and Jack Robison}
\date{}

@@

\section{Introduction}
LBRY \citep{li2020lbry} needs a trending algorithm to rank claims by `recent popularity',
preferably using on-chain information such as changes in support. An
intuitive idea is to have increases and decreases in overall support cause
a `spike' in the trending score, which then fades away over time
(assuming no further changes occur). This idea, which corresponds to
AR(1) time series in statistics \citep{brockwell2009time}
has formed the basis of all the implementations I have written --- though
other implementations have also existed in the past.

@@

In 2021, Jack Robison rewrote a lot of the hub code to improve performance
and to ensure hub data integrity when a blockchain reorganisation occurs. The
backend now uses leveldb to store transaction and claim information, and
Elasticsearch to store claim information in a suitable form for clients to
query. Leveldb will be replaced with rocksdb soon.
As part of these changes, trending needed to be made more efficient. This
meant we made some modifications to it, some of which are pretty neat
in my opinion.

@@

{\bf TODO}: Consider idea of symmetrising by using both old and new LBC values,
as this may make reorg safety occur naturally without any extra effort

{\bf TODO}: Show the response of the trending score to different patterns
of support.

{\bf TODO}: Bring time delays back in, and write about why.

@@

\section{The `Squashed' Grade}
Suppose $x$ is a huge number, where double precision overflow might occur if
we were to use it in a calculation (this happens at around $10^{308}$).
If $x$ is positive, we could instead deal with logarithms. Then, to add
two huge numbers without overflow, we could use
\begin{align}
\log(x + y)  &= \log(\exp(\log x) + \exp(\log y)) \\
             &= \logsumexp(\log x, \log y),
\end{align}
where $\logsumexp()$ is given by
\begin{align}
\logsumexp(a, b) &= \max(a, b)
        + \log\left[\exp(a - \max(a, b)) + \exp(b - \max(a, b))\right].
\end{align}
Subtracting the maximum before doing the exponentiation prevents overflow.
Similarly, differences could be computed on the log scale. Assuming $x>y$,
we can use the following method:
\begin{align}
\log(x - y)  &= \log(\exp(\log x) - \exp(\log y)) \\
             &= \logdiffexp(\log x, \log y),
\end{align}
where $\logdiffexp$ is given by
\begin{align}
\logdiffexp(a, b) &= a
        + \log\left[1 - \exp(b-a)\right]
\end{align}
for $a > b$.
If we were to use logarithms, all orderings would remain the same, since
log is a monotonic function \citep{knuth2019lattices}.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.7\textwidth]{figures/response.pdf}
\caption{An example of how the trending score might vary over time
for a single claim. The green curve shows $Y(t)$ which illustrates
the `exponentially decaying spikes' concept. The orange curve
shows $Z(t)$ which uses the sparse representation, so the trending
score remains constant whenever nothing happens to the claim.
Finally, the blue curve shows $S(Z(t))$, after the squashing
function is applied to prevent overflow. This is what the hub server
actually computes.\label{fig:response}}
\end{figure}

@@

\section{Acknowledgements}
BJB would like to thank Kevin Knuth (SUNY Albany) for inspiration, and the
LBRY Inc staff for building such a cool thing.
